### 2019-03-06 paper/sex_selection_spacing-ver2.tex
Update 1st paragraph in Introduction

Birth spacing has long served as a measure of son preference, with strong
son preference typically associated with shorter birth intervals the
fewer sons a family has \citep{ben-porath76b,Leung1988}.
However, the introduction of prenatal sex determination has the potential to
fundamentally change the relationship between son preference and birth spacing.
The stronger the son preference, the more likely couples are to resort to sex-selective 
abortions, and each abortion automatically increases the interval between births by six 
months to a year.%
\footnote{
The increase consists of three parts.
First, starting from the time of the abortion, the uterus needs at 
least two menstrual cycles to recover;  otherwise, the likelihood 
of spontaneous abortion increases substantially \citep{zhou00b}.
The second part is the waiting time to conception, which is between 
one and six months \citep{Wang2003}.
Finally, sex determination tests are reliable only from three months 
of gestation onwards.
}
As a result, we may now observe \emph{longer} average spacing for families with the
strongest son preference, precisely because it makes them more likely to use sex selection.
% To further complicate matters, families with son preference who---for one reason or 
% another---do not use prenatal sex selection may continue to have shorter birth spacing 
% than those with a lower preference for sons.

@@

Second, birth spacing can have a substantial impact on children's outcomes.
Most of the research finds a negative impact of short spacing on child health, although 
identifying the causal mechanisms have proven more difficult
\citep{Conde-Agudelo2006,Conde-Agudelo2012}.
There are three pathways that birth intervals can affect health.
First, there is maternal depletion, where the mother does not have sufficient time to 
recover after her last pregnancy before she become pregnant again.%
\footnote{
This depletion can take the form of nutrition, folate, or cervical insufficiency.
}
Second, disease transmission both from mother to child and between children close in age.
Third, sibling competition, where siblings close in age compete for scarce resources and
parental care.
Which mechanism dominates is of particular interest when son preference drives spacing 
decisions.
Both the maternal depletion and the transmission of disease explanations predominately
affect the subsequent child and the mother.
Only sibling competition has the potential to negatively affect the older child.

@@

\input{../figures/appendix_spell2_low.tex}

@@

\input{../figures/appendix_spell2_med.tex}

@@

\input{../figures/appendix_spell2_high.tex}

@@

% low education
\input{../figures/appendix_spell3_low.tex}

@@

\input{../figures/appendix_spell3_med.tex}

@@

\input{../figures/appendix_spell3_high.tex}

@@

\input{../figures/appendix_spell4_low.tex}

@@

\input{../figures/appendix_spell4_med.tex}

@@

\input{../figures/appendix_spell4_high.tex}

>>>>

Birth spacing has long served as a measure of son preference, with strong
son preference typically associated with shorter birth intervals the
fewer sons a family has \citep{ben-porath76b,Leung1988}.
However, the introduction of prenatal sex determination has the potential to
fundamentally change the relationship between son preference and birth spacing.
The stronger the son preference, the more likely couples are to resort to sex-selective 
abortions, and each abortion automatically increases the interval between births by six 
months to a year.%
\footnote{
The increase consists of three parts.
First, starting from the time of the abortion, the uterus needs at 
least two menstrual cycles to recover;  otherwise, the likelihood 
of spontaneous abortion increases substantially \citep{zhou00b}.
The second part is the waiting time to conception, which is between 
one and six months \citep{Wang2003}.
Finally, sex determination tests are reliable only from three months 
of gestation onwards.
}
Longer birth intervals may, therefore, no longer be an indication of declining son
preference, but could instead come about precisely because strong son preference makes
couples more likely to use sex selection.
% To further complicate matters, families with son preference who---for one reason or 
% another---do not use prenatal sex selection may continue to have shorter birth spacing 
% than those with a lower preference for sons.

@@

Second, birth spacing can have a substantial impact on children's outcomes.
Most of the research finds a negative impact of short spacing on child health, although 
identifying the causal mechanisms have proven more difficult
\citep{Conde-Agudelo2006,Conde-Agudelo2012}.
There are three pathways that birth intervals can affect health.
First, there is maternal depletion, where the mother does not have sufficient time to 
recover after her last pregnancy before she becomes pregnant again.%
\footnote{
This depletion can take the form of nutrition, folate, or cervical insufficiency.
}
Second, disease transmission both from mother to child and between children close in age.
Third, sibling competition, where siblings close in age compete for scarce resources and
parental care.
Which mechanism dominates is of particular interest when son preference drives spacing 
decisions.
Both the maternal depletion and the transmission of disease explanations predominately
affect the subsequent child and the mother.
Only sibling competition has the potential to negatively affect the older child.

@@

% \input{../figures/appendix_spell2_low.tex}

@@

% \input{../figures/appendix_spell2_med.tex}

@@

% \input{../figures/appendix_spell2_high.tex}

@@

% low education
% \input{../figures/appendix_spell3_low.tex}

@@

% \input{../figures/appendix_spell3_med.tex}

@@

% \input{../figures/appendix_spell3_high.tex}

@@

% \input{../figures/appendix_spell4_low.tex}

@@

% \input{../figures/appendix_spell4_med.tex}

@@

% \input{../figures/appendix_spell4_high.tex}
